\textls[-0]{\thetitle}

### 2023-03-14 metatheory-condensed.tex
add comments from Slack

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
(and \cref{fig:fmltt-selected} \TT).
Dependent function types \nTyPi{x}{A}{B},
dependent pair types \nTySigma{x}{A}{B}, and
identity types \TyId{t_1}{t_2} are standard.
So are their introduction and elimination forms.
Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement. It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\ 
Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation. Similar for ( _ |- ) vs. (G |-) and (G |- _ ) vs (G |- T). \\ 
I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e. there are no two different derivations lead to the same t; for each t, there is only one derivation for it). This is not true for extrinsic typing. \\
If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
}
we introduce judgmental equality (of the form \goodTerm{\Gm}{t_1 \equiv t_2}{T})
indicating convertibility.

>>>>

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
(and \cref{fig:fmltt-selected} \TT).
Dependent function types \nTyPi{x}{A}{B},
dependent pair types \nTySigma{x}{A}{B}, and
identity types \TyId{t_1}{t_2} are standard.
So are their introduction and elimination forms.
Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
Definitional equalities have the form
\goodTerm{\Gm}{t_1 \equiv t_2}{T} and \goodType{\Gm}{T_1 \equiv T_2}{}.\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B. Maybe this is something require emphasis? Because it is usually handled by LF (like QIIT) but now we are using bare hands so.... maybe we need to emphasize this}
%For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement. It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\ 
%Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation. Similar for ( _ |- ) vs. (G |-) and (G |- _ ) vs (G |- T). \\ 
%I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e. there are no two different derivations lead to the same t; for each t, there is only one derivation for it). This is not true for extrinsic typing. \\
%If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
%I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
%}
